---
title: "<div class='mytitle'>Benchmarking</div>"
#author: "<center>[Christoph Ogris](https://github.com/cellmapslab/kimono) </center>"
date: "<center>`r format(Sys.time(), '%d %B %Y')`</center>"
mail: "christoph.ogris@helmholtz-muenchen.de"
linkedin: "christoph-ogris-6a223082"
twitter: "christoph_ogris"
github: "00chris00"
home: "www.helmholtz-muenchen.de/icb/index.html"
logo: "logo.png"
output:
  lazyrmd::lazy_render:
    toc: TRUE
    toc_float:
      collapsed: true
      smooth_scroll: false
    number_sections: FALSE
    css: style.css
    self_contained: false
---

# Libraries
```{r, warning = FALSE,  message=FALSE}
library(kimono) # framework v 2.0.1.0006
```

## Load Data

```{r, warning = FALSE}
#load('data/220221_benchmark.RData')
load('data/220221_benchmark.RData')
```

## Remove Phenotype Features

```{r, warning = FALSE}
#input_data$phenotype <- input_data$phenotype[,-c('srv.time','srv.status')]
```


## DEBUG EXAMPLE

```{r, warning = FALSE}
DEBUG_nodes <- V(prior_network)[name %like% c('MAPK1','MYC')]
DEBUG_subnet <- ego(prior_network, order=1, nodes = DEBUG_nodes, mode = "all", mindist = 0)
DEBUG_prior_network <- induced_subgraph(prior_network,unlist(DEBUG_subnet))
rm(DEBUG_nodes, DEBUG_subnet)
```

#### Plotting prior network
```{r, warning = FALSE}
vertex <- do.call(rbind,strsplit(V(DEBUG_prior_network)$name,split = '___'))

DEBUG_prior_network %>% plot(edge.curved=0,
                       main = 'DEBUG Prior Network',
     vertex.color = c("steel blue", "orange","limegreen","grey")[vertex[,1] %>% as.factor %>% as.numeric],
     vertex.frame.color="white",
     vertex.label = vertex[,2], 
     vertex.label.color='black',
     vertex.label.cex=.7,
     layout=layout_randomly, rescale=F) 
```

# Dealing with missing data 
## Multivariate Imputation by Chained Equations
The fully observed variables are cnv and mutation. Partially observed variables are genes, methylation, phenotype, time, status.

```{r, warning=TRUE}
library(impute)

# Which data is missing?
message(paste0("# of missing values by genes ",sum(is.na(input_data$gene))))
message(paste0("# of missing values by methylation ",sum(is.na(input_data$methylation))))
message(paste0("# of missing values by cnv ",sum(is.na(input_data$cnv))))
message(paste0("# of missing values by phenotype overall ",sum(is.na(input_data$phenotype))))
message(paste0("# of missing values by phenotype: time ",sum(is.na(input_data$phenotype$srv.time))))
message(paste0("# of missing values by phenotype: status ",sum(is.na(input_data$phenotype$srv.status))))
message(paste0("# of missing values by mutation ",sum(is.na(input_data$mutation))))

THRESHOLD_MISSING <- 0.2
SEED <- 42
K <- 10

# GENES
# rows are sample IDs, columns are gene IDs
gene <- as.data.frame(input_data$gene)
# transform
gene.t <- t(gene)
print(dim(gene.t))
message("genes x samples before the knn imputation ")
geneImp <- impute.knn(gene.t, k=K, rowmax = THRESHOLD_MISSING, colmax = THRESHOLD_MISSING, rng.seed = SEED)
imputedGene <- t(geneImp$data)
message(paste0("# of missing values by genes after imputation ",sum(is.na(imputedGene))))

#METHYLATION
# rows are sample IDs, columns are gene methylation sites
methylation <- as.data.frame(input_data$methylation)
# transform
methylation.t <- t(methylation)
print(dim(methylation.t))
message("methylation x samples before the knn imputation")
methylationImp <- impute.knn(methylation.t, k=K, rowmax = THRESHOLD_MISSING, colmax = THRESHOLD_MISSING, rng.seed = SEED)
imputedMethylation <- t(methylationImp$data)
message(paste0("# of missing values by methylation after imputation ",sum(is.na(imputedMethylation))))

imputed_data <- input_data
imputed_data$gene <- as.data.table(imputedGene)
imputed_data$methylation <- as.data.table(imputedMethylation)

message(paste0("# of missing values by genes ",sum(is.na(imputed_data$gene))))
message(paste0("# of missing values by methylation ",sum(is.na(imputed_data$methylation))))
message(paste0("# of missing values by cnv ",sum(is.na(imputed_data$cnv))))
message(paste0("# of missing values by phenotype overall ",sum(is.na(imputed_data$phenotype))))
message(paste0("# of missing values by phenotype: time ",sum(is.na(imputed_data$phenotype$srv.time))))
message(paste0("# of missing values by phenotype: status ",sum(is.na(imputed_data$phenotype$srv.status))))
message(paste0("# of missing values by mutation ",sum(is.na(imputed_data$mutation))))
```

# KiMONo

## Call KiMONo
### Input data
```{r, warning = FALSE}
network <- kimono(input_data, DEBUG_prior_network, core = 2, infer_missing_prior = TRUE)
#plot network
to_igraph(network) %>% plot_kimono(title='KiMONo Network (directed)')
to_igraph(network, directed = F) %>% plot_kimono(title='KiMONo Network (undirected)')
```
### Imputed data
```{r, warning = FALSE}
network_imputed <- kimono(imputed_data, DEBUG_prior_network, core = 2, infer_missing_prior = TRUE)
#plot network
to_igraph(network_imputed) %>% plot_kimono(title='KiMONo imputed Network (directed)')
to_igraph(network_imputed, directed = F) %>% plot_kimono(title='KiMONo imputed Network (undirected)')
```

## Result
<br>
Columns:

  * **target** - vector y in regression model
  * **predictor** - each feature in X used in model y
  * **value** - effect size of predictor on target
  * **r_squared** - model performance
  * **mse** - model error
  * **predictor_layer** - input data the predictor belongs to
  * **target_layer** - input data the target belongs to

<br>
```{r, warning = FALSE}
DT::datatable(head(network), class = 'cell-border stripe')
```
```{r, warning = FALSE}
DT::datatable(head(network_imputed), class = 'cell-border stripe')
```

# Network Analysis
***

## Quality

Evaluating the r2 for each model gives us the possibility to compare the performances of our models.
### Input data
> Note the rsquared is the same for each target.  

```{r, warning = FALSE}
gg_all <- network[predictor == '(Intercept)',] %>%  
            ggplot( aes(y=r_squared))  +
                    geom_boxplot()

gg_grouped <- network[predictor == '(Intercept)',] %>%  
            ggplot( aes(y=r_squared,x=target_layer))  +   
                    geom_boxplot(fill=c("steel blue",'#842F39', "orange"))

nnodes <- c(network$target, network$predictor) %>% unique %>% length
nedges <- dim(network)[1]

cat('Number of Nodes: ',nnodes)
cat('Number of Edges: ',nedges)
```
### Imputed data
```{r, warning = FALSE}
gg_all_imputed <- network_imputed[predictor == '(Intercept)',] %>%  
            ggplot( aes(y=r_squared))  +
                    geom_boxplot()

gg_grouped_imputed <- network_imputed[predictor == '(Intercept)',] %>%  
            ggplot( aes(y=r_squared,x=target_layer))  +   
                    geom_boxplot(fill=c("steel blue",'#842F39', "orange"))

nnodes_imputed <- c(network_imputed$target, network_imputed$predictor) %>% unique %>% length
nedges_imputed <- dim(network_imputed)[1]

cat('Number of Nodes: ',nnodes_imputed)
cat('Number of Edges: ',nedges_imputed)
```

## Filter network 

Often we are only interested in models which perform well and have large effect sizes.
### Input data
```{r, warning = FALSE}
network <- network %>% 
  filter(value > 0.001 | value < -0.001 ) %>% # filter low effects
  filter(r_squared > 0.001)  %>%  # filter low performing models
  filter(predictor != '(Intercept)') # filter all intercepts (should be close to 0 due to normalization step)
```

New network properties
```{r, warning = FALSE}
nnodes <- c(network$target, network$predictor) %>% unique %>% length
nedges <- dim(network)[1]

cat('Number of Nodes: ',nnodes)
cat('Number of Edges: ',nedges)
```

Generate igraph for easier network analysis
```{r, warning = FALSE}
#generate undirected igraph
ig_network <- to_igraph(network, directed=TRUE) 
ig_network %>% plot_kimono
```

## General
```{r, warning = FALSE}
cat('Density: ', ecount(ig_network)/(vcount(ig_network)*(vcount(ig_network)-1)), '\n' )
cat('Reciprocity: ', reciprocity(ig_network) , '\n' )
cat('Transitivity: ', transitivity(as.undirected(ig_network, mode="collapse")) , '\n' )
```
### Imputed data
```{r, warning = FALSE}
network_imputed <- network_imputed %>% 
  filter(value > 0.001 | value < -0.001 ) %>% # filter low effects
  filter(r_squared > 0.001)  %>%  # filter low performing models
  filter(predictor != '(Intercept)') # filter all intercepts (should be close to 0 due to normalization step)
```

New network_imputed properties
```{r, warning = FALSE}
nnodes <- c(network_imputed$target, network_imputed$predictor) %>% unique %>% length
nedges <- dim(network_imputed)[1]

cat('Number of Nodes: ',nnodes)
cat('Number of Edges: ',nedges)
```

Generate igraph for easier network_imputed analysis
```{r, warning = FALSE}
#generate undirected igraph
ig_network_imputed <- to_igraph(network_imputed, directed=TRUE) 
ig_network_imputed %>% plot_kimono
```

## General
```{r, warning = FALSE}
cat('Density: ', ecount(ig_network_imputed)/(vcount(ig_network_imputed)*(vcount(ig_network_imputed)-1)), '\n' )
cat('Reciprocity: ', reciprocity(ig_network_imputed) , '\n' )
cat('Transitivity: ', transitivity(as.undirected(ig_network_imputed, mode="collapse")) , '\n' )
```
# Session
***
```{r , eval = FALSE }
 sessionInfo()
```


